---
title: "Machine Learning Course Project"
author: "dmontiel242"
date: "October 25, 2015"
output: html_document
---

## Synopsis: 
A random forest model with cross-validation is applied to fitness data in order to assess activity quality, which is the factor variable classe in the pml data set. The model creates a training and test set using the createDataPartition function in the caret package. 

## Out of Sample Error and cross-validation

The error rate from the test set provides a measure of the out of sample error. Cross-validation is performed using the trControl option in the train in the caret package and is set to use k-folds cross-validation with 5 folds.

## Data Pre-Processing:
The data is preprocessed using principal component analysis , variables with near zero variance are removed and all remaining variables are converted to numeric variables. 


```{r}
library(caret)
library(randomForest)
url <- 'https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv'

if (!exists('pmldata')){
  print('dowloading data')
  pml <- download.file(url,destfile = "C:\\Users\\Dan\\Desktop\\Coursera\\pmltraining.csv")
  pmldata = read.csv("C:\\Users\\Dan\\Desktop\\Coursera\\pmltraining.csv")
}
#subset data into training and test sets for cross-validation
#inTrain <- createDataPartition(y=pmldata$classe, p = 0.75,list = FALSE)
classy <- factor(c(pmldata$classe))
numcol <- length(colnames(pmldata))
#isnumer <- sapply(pmldata, is.numeric)
pmlnumer <- pmldata[,-numcol]
pmlnumer <-pmldata
pmlnumer <- pmlnumer[,-1]
pmlnumer <- pmlnumer[,-1]
pmlnumer <- pmlnumer[,-1]
pmlnumer <- pmlnumer[,-1]
pmlnumer <- pmlnumer[,-1]
pmlnumer <- pmlnumer[,-1]
pmlnumer <- pmlnumer[,-1]
pmlnumer <- data.frame(sapply(pmlnumer, as.numeric))
pmldat<- pmlnumer
#pmldat <- pmldat[complete.cases(pmldat),]
nzv <- nearZeroVar(pmldat)
pmldat <- pmldat[,-nzv]
row.names(pmldat) <- 1:nrow(pmldat)
pmldat[is.na(pmldat)]<-0
pmldat[,'classe'] <- factor(pmldat[,'classe'])
inTrain <- createDataPartition(y=pmldat$classe, p = 0.7,list = FALSE)
training <- pmldat[inTrain,]
testing <- pmldat[-inTrain,]
preProc <- preProcess(training, method = 'pca', thresh = 0.90)
trainPC <- predict(preProc,training)
```
The model was applied with the train function in the caret package but was not executed to save time in knitting this document.

modelfit2 <- train(training$classe ~., method = 'rf',trControl=trainControl(method="cv",number=5),data=trainPC)
saveRDS(modelfit2,'mysavedModel002.rds')


Using submission set was downloaded and pre-processed in the same way as the training set. The predict function was applied to the cleaned and processed submission set with the same PCA preprocessing parameters.
```{r}

library(RCurl)

url2 <- 'https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv'

if (!exists('testdata')){
testdat <- getURL(url2,
               ssl.verifypeer=0L, followlocation=1L)

#  testdat <- download.file(url2,destfile = "C:\\Users\\Dan\\Desktop\\Coursera\\testdata.csv")
testdata <-  read.csv("C:\\Users\\Dan\\Desktop\\Coursera\\testdata.csv")
}
testdata<- data.frame(sapply(testdata, as.numeric))
testdata <- testdata[,-1]
testdata <- testdata[,-1]
testdata <- testdata[,-1]
testdata <- testdata[,-1]
testdata <- testdata[,-1]
testdata <- testdata[,-1]
testdata <- testdata[,-1]

testdata <- testdata[,-nzv]
testdata[is.na(testdata)]<-0

#testpc <- predict(preProc,testdata)
#answers = factor(predict(modelfit2,testpc),labels=c('A','B','C','D','E'))
pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}
#pml_write_files(answers)
```

